---
title: "Algatr"
format: pdf
editor: visual
---

-   algatr help website: <https://thewanglab.github.io/algatr/>

-   algatr github: <https://github.com/TheWangLab/algatr_workshop>

```{r}

library(installr) 
install.packages.zip("https://cran.csiro.au/bin/macosx/big-sur-x86_64/contrib/4.5/wingen_2.2.0.tgz")


cp -R -i -n /Library/Frameworks/R.framework/Versions/4.3-x86_64/Resources/library/ /Library/Frameworks/R.framework/Resources/library
```

\^ for installing packages using zip file from cran Installing it just using base r `install.packages` (wasn't working for some reason, it was returning a 404 error page)

## Algatr Tutorial

```{r}
library(algatr)
# alazygatr_packages()
    # ^updates packages

#Warning message:
#In i.p(...) :
#  installation of package ‘/var/folders/5g/bpbtlbnn6n9_ptfrqwyy_wkh0000gn/T//RtmpsX4vU0/filea4c36bd44d75/RStoolbox_1.0.2.1.tar.gz’ had non-zero exit status

load_algatr_example()
plot(CA_env, col = turbo(100), axes = FALSE)

env <- scaleRGB(CA_env)
plotRGB(env, r = 1, g = 2, b = 3)

# Add sampling localities on top of this
points(liz_coords, pch = 19)

```

### Files needed:

-   `do_everything_for_me(liz_vcf, liz_coords, CA_env)`

-   vcf file, coordinate file (long=x, lat=y), and env file

env file

-   rasterstack object (<https://rspatial.github.io/raster/reference/stack.html>)

-   Example env file

    -   These were generated by performing a raster PCA on 19 bioclimatic variables (obtained from the World Clim database) and retaining the top 3 PCs

    -   We can combine all three PCs into a single map by scaling each of the rasters such that they each correspond to either R, G, or B using the [`scaleRGB()`](https://thewanglab.github.io/algatr/reference/scaleRGB.html) function and subsequently map using the `plotRGB()` function.

        ```{r}
        env <- scaleRGB(CA_env)
        plotRGB(env, r = 1, g = 2, b = 3)
        ```

### Getting Raster data into R

Tutorials:

-   <https://wec.wur.nl/r/spatial/raster-data.html>

-   <https://ouicodedata.com/posts/worldclim-with-r/>

```{r}
library(raster)
library(sp)
library(geodata)
library(algatr)

raster_path = "/Users/caprinapugliese/Documents/School/Uconn/2024-26_Grad_School/Dagilis-lab/WNS-project/data/climate_data/wc2.1_data"

bio <- worldclim_global(path=raster_path, var="bio", res=2.5)

# information on the coordinate system
crs(CAN_bio, proj = TRUE, describe = TRUE)
crs(USA_bio, proj = TRUE, describe = TRUE)

plot(bio[[12]])
plot(CAN_bio[[12]])

# crop Raster* with Spatial* object
NoA <- as(extent(-135, -55, 25, 60), 'SpatialPolygons')
crs(NoA) <- crs(bio)
NoA_bio <- crop(bio, NoA)
plot(NoA_bio[[12]])

## https://stackoverflow.com/questions/75739085/running-a-pca-on-a-rasterstack-in-r
## pca code
pca <- prcomp(NoA_bio)
x <- predict(NoA_bio, pca)
plot(x)

NoA_pca_map <- plotRGB(scaleRGB(x),r=1,g=2,b=3)
```

```{r}
# example using multiple countries
w <-  worldclim_country(c("BEL", "NL", "Luxembourg"), var="tavg")
m <- merge(w)
plot(bio_m, 1)

# plot the 12 bio element
plot(bio[[12]])


library(ggplot2)
ggplot(bio) +
  geom_sf(fill = "#69b3a2", color = "white") +
  theme_void()

library(leaflet)
map <- addTiles(leaflet())
# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/backgroundMapBasic.html"))

map 
```

-   data

    -   x –\> long min: -121.5, max: -62

    -   y –\> lat min: 34.5, max: 48.5